---
title: "Watch Market Analysis for Henry Wilson Jewelers"
author: "Brian Waite"
date: "January 3, 2019"
output:
  prettydoc::html_pretty:
    theme: tactile
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


```{r include = FALSE}
#Set up analysis environment
library(tidyr)
library(ggplot2)
library(ggthemes)
library(rvest)
library(stringr)
library(reshape2)
```

### Background: 

Every fall *National Jeweler* magazine publishes a special edition titled, *The State of the Majors*. The magazine self-describes this as "an analysis of the jewelry industry's high-volume sellers."

These are available online here: 

https://www.nationaljeweler.com/magazine/stateofthemajors

Sellers are grouped into two primary categories and may be found in both: 

1. **Supersellers** -- Retailers that sold $100 Million+ in jewelry and watches in North America in their latest fiscal year. 

2. **Top 50 Specialty Jewelers** -- The largest jewelry chains in North America ranked by store count. 

For each store, a number of characteristics are printed including store names, number of stores, key executives and website information. Most importantly, the rankings list key brands for each of the retailers. In most cases watches are listed separately from jewelry items. 

<br /> 

### Purpose of this analysis: 

Using the information provided by the National Jeweler publication, this analysis hopes to explore whether insights can be gained regarding watch brand market performance, what brands the top retailers in the country are carrying, and potentially any trends in the popularity of watch brands over time. 

Ultimately, this research hopes to support informed decision making regarding future inventory. 

It is important to note the data below is not *sales* data, but rather is gathered from a list of **key brands** as identified by the retailer. 

### 2018 Analysis: 

```{r}
#Base URL for National Jeweler State of the Majors Report

baseurl = ("https://www.nationaljeweler.com/media/stateofthemajors/2018/inc/html/")


#Get data for top 50 specialty jewelry stores

specialtyURL = c("34","35","36","38","39","40","42","43","44","46")
specialtyData = data.frame()

for (items in specialtyURL) { 
  url = paste(baseurl,items,".html",sep = "")
  site = read_html(url)
  job = site %>%
  html_nodes("div") %>%
  html_text()
  job = job[2]
  outcome = str_split(job,"[:digit:]{1}\\.\\s|\\(tie\\)", simplify = TRUE)
  specialtyData = rbind(specialtyData,t(outcome))
}

#supersellers data -- retailers that sold $100M+ in Jewelry and Watches in North America in latest FY

supersellersURL = c("20","21","22","24","26","27","28","30","31","32")
supersellerData = data.frame()

for (items in supersellersURL) { 
  url = paste(baseurl,items,".html",sep = "")
  site = read_html(url)
  job = site %>%
    html_nodes("div") %>%
    html_text()
  job = job[2]
  outcome = str_split(job,"[:digit:]{1}\\.\\s|\\(tie\\)", simplify = TRUE)
  supersellerData = rbind(supersellerData,t(outcome))
}

# Get company names and Watches sold separated out

#supersellers

supersellerList = list()
supersellerWatches = list()

for (stores in supersellerData$V1) {
supersellerList = paste(supersellerList,str_split(stores,"[:digit:]")[[1]][1],sep = ",")
supersellerWatches = paste(supersellerWatches, str_split(stores,".*Watches:\\s*|\\. WEB.*")[[1]][2], sep = ";")
}

supersellerList = as.list(str_split(supersellerList,","))[[1]]
supersellerWatches = as.list(str_split(supersellerWatches,";"))[[1]]
supersellerWatches[27] = ""

supersellerList = supersellerList[-c(1,2,5,8:32,36,41,47,48,53,57,60,63,64,65,69,72,75,81,84,85)]
supersellerWatches = supersellerWatches[!supersellerWatches=="NA"]
supersellerWatches = supersellerWatches[-c(1)]

#Specialty

specialtyList = list()
specialtyWatches = list()

for (stores in specialtyData$V1) {
  specialtyList = paste(specialtyList,str_split(stores,"[:digit:]")[[1]][1],sep = ",")
  specialtyWatches = paste(specialtyWatches, str_split(stores,".*Watches:\\s*|\\. WEB.*")[[1]][2], sep = ";")
}

specialtyList = as.list(str_split(specialtyList,","))[[1]]
specialtyWatches = as.list(str_split(specialtyWatches,";"))[[1]]

specialtyList = specialtyList[-c(1,2,6,12,18,25,29,31:33,35,38,42,50,53,58,60,67)]
specialtyWatches = specialtyWatches[-c(1)]
specialtyWatches = specialtyWatches[!specialtyWatches=="NA"]


# Now merge the two lists

supersellers = as.data.frame(cbind(supersellerList,supersellerWatches),stringsAsFactors = FALSE)
supersellers$supersellerWatches = as.list(sapply(supersellers$supersellerWatches, str_split,","))
supersellers = unnest(supersellers)
supersellers[which(supersellers$supersellerWatches == "IWC"),2] = "IWC Schaffhausen"
supersellers$supersellerWatches = trimws(supersellers$supersellerWatches)
supersellers = rbind(supersellers, c("AMAZON.COM INC.","Casio"))
supersellers[87,2] = "Bulova"
supersellers$supersellerWatches = trimws(supersellers$supersellerWatches)

#Lots of cleanup

specialty = as.data.frame(cbind(specialtyList, specialtyWatches),stringsAsFactors = FALSE)
specialty$specialtyWatches = as.list(sapply(specialty$specialtyWatches, str_split,","))
specialty = unnest(specialty)
specialty[which(specialty$specialtyWatches == "IWC"),2] = "IWC Schaffhausen"
specialty$specialtyWatches = trimws(specialty$specialtyWatches)
specialty[54,2] = "Tissot"
specialty$specialtyWatches = trimws(specialty$specialtyWatches)
specialty[which(specialty$specialtyWatches == "IWC Schauffhausen"),]$specialtyWatches = "IWC Schaffhausen"
specialty[which(specialty$specialtyWatches == "IWC"),]$specialtyWatches = "IWC Schaffhausen"
specialty[which(specialty$specialtyWatches == "Jaeger LeCoultre"),]$specialtyWatches = "Jaeger-LeCoultre"
specialty[which(specialty$specialtyWatches == "A. Lange and Sohne"),]$specialtyWatches = "A. Lange & Söhne"
specialty[which(specialty$specialtyWatches == "Tag Heuer"),]$specialtyWatches = "TAG Heuer"
specialty[which(specialty$specialtyWatches == "Frederique Constant"),]$specialtyWatches = "Frédérique Constant"
specialty[which(specialty$specialtyWatches == "Baume and Mercier"),]$specialtyWatches = "Baume & Mercier"
specialty[which(specialty$specialtyWatches == "Dior Watches"),]$specialtyWatches = "Dior"
specialty[which(specialty$specialtyWatches == "G-Shock by Casio"),]$specialtyWatches = "G-Shock"
specialty[which(specialty$specialtyWatches == "Baby G"),]$specialtyWatches = "Baby-G"

#Get the number of times each brand appears in the above list
topWatch1 = table(supersellers$supersellerWatches)
topWatch1 = as.data.frame(topWatch1[-c(1)])
topWatch1 = topWatch1[order(-topWatch1$Freq),]
names(topWatch1) = c("Brand","Freq")

#Get the number of times each brand appears in the above list
topWatch2 = table(specialty$specialtyWatches)
topWatch2 = as.data.frame(topWatch2[-c(1)])
topWatch2 = topWatch2[order(-topWatch2$Freq),]
names(topWatch2) = c("Brand","Freq")

```

<br /> 

#### 2018 Supersellers: 

Although these retailers represent the largest in North America, understanding watch brands they carry may provide some useful insight. First, we can examine which retailers carry the most brands so we have some context: 

```{r}
topSuperSellers = table(supersellers$supersellerList)
topSuperSellers = as.data.frame(topSuperSellers[-c(1)])
topSuperSellers = topSuperSellers[order(-topSuperSellers$Freq),]
names(topSuperSellers) = c("Retailer","Number of Watch Brands Carried")
knitr::kable(topSuperSellers,row.names = FALSE)
```

The below plot shows the top 30 watch brands as ranked by the number of superseller retailers who carry them. 

```{r fig.height = 7, fig.width = 11}
p1 = ggplot(data = topWatch1[1:30,], aes(x = reorder(Brand,Freq), y = Freq)) + 
  geom_bar(stat = "identity") + ggtitle("Number of 2018 Supersellers\nIdentifying as Key Brand (Top 30)") +
  ylab("Number of Retailers Carrying Brand") + 
  xlab("Brand") 
p1 + coord_flip() + theme_wsj()
```

<br/>

#### 2018 Top 50 Specialty Retailers: 

The same information from Specialty Jewlery store chains may be more interesting as the size of individual stores more closely reflects the market segment for Henry Wilson Jewelers. These retailers include: 

```{r}
topSpecialty = table(specialty$specialtyList)
topSpecialty = as.data.frame(topSpecialty[-c(1)])
topSpecialty = topSpecialty[order(-topSpecialty$Freq),]
names(topSpecialty) = c("Retailer","Number of Watch Brands Carried")
knitr::kable(topSpecialty,row.names = FALSE)
```
<br /> 

```{r fig.height = 7, fig.width = 11}
p2 = ggplot(data = topWatch2[1:30,], aes(x = reorder(Brand,Freq), y = Freq)) + 
  geom_bar(stat = "identity") + ggtitle("Number of 2018 Specialty Retailers\n(Top 50 by Num. Stores)\nIdentifying as Key Brand (Top 30)") +
  ylab("Number of Retailers Carrying Brand") + 
  xlab("Brand") 
p2 + coord_flip() + theme_wsj()
```

<br /> 

#### Analysis of Data Over Time

In addition to the 2018 magazine, data is also available beginning in 2016 allowing to observe how the brands specific retailers have carried over time have changed. For the purposes of this historical analysis, the focus will be solely on Specialty Jewelers and not the Supersellers. 

```{r}
# Historical Data

# 2017 Specialty Data
baseurl = ("https://www.nationaljeweler.com/media/stateofthemajors/2017/inc/html/")

specialtyURL2017 = c("32","33","34","36","38","40","41","42","44","46")
specialtyData2017 = data.frame()

for (items in specialtyURL2017) { 
  url = paste(baseurl,items,".html",sep = "")
  site = read_html(url)
  job = site %>%
    html_nodes("div") %>%
    html_text()
  job = job[2]
  outcome = str_split(job,"[:digit:]{1}\\.\\s|\\(TIE\\)", simplify = TRUE)
  specialtyData2017 = rbind(specialtyData2017,t(outcome))
}

#Specialty 2017

specialtyList2017 = list()
specialtyWatches2017 = list()

for (stores in specialtyData2017$V1) {
  specialtyList2017 = paste(specialtyList2017,str_split(stores,"[:digit:]")[[1]][1],sep = ",")
  specialtyWatches2017 = paste(specialtyWatches2017, str_split(stores,".*Watches:\\s*|\\. WEB.*")[[1]][2], sep = ";")
}

specialtyList2017 = as.list(str_split(specialtyList2017,","))[[1]]
specialtyWatches2017 = as.list(str_split(specialtyWatches2017,";"))[[1]]

specialtyList2017 = specialtyList2017[-c(1,2,6,12,19,23,27,29,34,41,45,49,53,55)]
specialtyWatches2017 = specialtyWatches2017[-c(1)]
specialtyWatches2017 = specialtyWatches2017[!specialtyWatches2017=="NA"]

specialty2017 = as.data.frame(cbind(specialtyList2017, specialtyWatches2017),stringsAsFactors = FALSE)
specialty2017$specialtyWatches2017 = as.list(sapply(specialty2017$specialtyWatches2017, str_split,","))
specialty2017 = unnest(specialty2017)
specialty2017$specialtyWatches2017 = trimws(specialty2017$specialtyWatches2017)
specialty2017[which(specialty2017$specialtyWatches2017 == "IWC Schauffhausen"),]$specialtyWatches2017 = "IWC Schaffhausen"
specialty2017[which(specialty2017$specialtyWatches2017 == "IWC"),]$specialtyWatches2017 = "IWC Schaffhausen"
specialty2017[which(specialty2017$specialtyWatches2017 == "Jaeger LeCoultre"),]$specialtyWatches2017 = "Jaeger-LeCoultre"
specialty2017[which(specialty2017$specialtyWatches2017 == "A. Lange and Sohne"),]$specialtyWatches2017 = "A. Lange & Söhne"
specialty2017[which(specialty2017$specialtyWatches2017 == "Tag Heuer"),]$specialtyWatches2017 = "TAG Heuer"
specialty2017[which(specialty2017$specialtyWatches2017 == "Bell& Ross"),]$specialtyWatches2017 = "Bell & Ross"
specialty2017[which(specialty2017$specialtyWatches2017 == "Frederique Constant"),]$specialtyWatches2017 = "Frédérique Constant"
specialty2017[which(specialty2017$specialtyWatches2017 == "Baume and Mercier"),]$specialtyWatches2017 = "Baume & Mercier"
specialty2017[which(specialty2017$specialtyWatches2017 == "Dior Watches"),]$specialtyWatches2017 = "Dior"
specialty2017[which(specialty2017$specialtyWatches2017 == "G-Shock by Casio"),]$specialtyWatches2017 = "G-Shock"
specialty2017[which(specialty2017$specialtyWatches2017 == "Tissot. Retail Store Brands: Tourbillon Boutique"),]$specialtyWatches2017 = "Tissot"

topWatch2017 = table(specialty2017$specialtyWatches2017)
topWatch2017 = as.data.frame(topWatch2017[-c(1)])
topWatch2017 = topWatch2017[order(-topWatch2017$Freq),]
names(topWatch2017) = c("Brand","Freq")


#Specialty 2016

specialty2016 = read.csv2("2016Specialty.txt")

names(specialty2016) = c("Combined")
specialty2016 = specialty2016 %>% separate(Combined, c("Stores","Brand"), ",",fill = "right", extra = "merge")
specialty2016$Brand = as.list(sapply(specialty2016$Brand, str_split,","))
specialty2016 = unnest(specialty2016)
specialty2016[which(specialty2016$Brand == " Company\\u2019s own brand."),]$Brand = "Company's own brand."
specialty2016$Brand = trimws(specialty2016$Brand)
specialty2016[which(specialty2016$Brand == "Tag Heuer"),]$Brand = "TAG Heuer"
specialty2016[which(specialty2016$Brand == "Victorinox Swiss Army"),]$Brand = "Victorinox"
specialty2016[which(specialty2016$Brand == "Victornox Swiss Army"),]$Brand = "Victorinox"
specialty2016[which(specialty2016$Brand == "A. Lange and Sohne"),]$Brand = "A. Lange & Söhne"
specialty2016[which(specialty2016$Brand == "A. Lange & S\\u00f6hne"),]$Brand = "A. Lange & Söhne"
specialty2016[which(specialty2016$Brand == "Baume and Mercier"),]$Brand = "Baume & Mercier"
specialty2016[which(specialty2016$Brand == "ESQ"),]$Brand = "ESQ by Movado"
specialty2016[which(specialty2016$Brand == "ESQ Movado"),]$Brand = "ESQ by Movado"
specialty2016[which(specialty2016$Brand == "Frederique Constant"),]$Brand = "Frédérique Constant"
specialty2016[which(specialty2016$Brand == "Fr\\u00e9d\\u00e9rique Constant"),]$Brand = "Frédérique Constant"
specialty2016[which(specialty2016$Brand == "IWC Schauffhausen"),]$Brand = "IWC Schaffhausen"
specialty2016[which(specialty2016$Brand == "IWC"),]$Brand = "IWC Schaffhausen"
specialty2016[which(specialty2016$Brand == "Jaeger LeCoultre"),]$Brand = "Jaeger-LeCoultre"
specialty2016[which(specialty2016$Brand == "Dior Watches"),]$Brand = "Dior"

topWatch2016 = table(specialty2016$Brand)
topWatch2016 = as.data.frame(topWatch2016[-c(1)])
topWatch2016 = topWatch2016[order(-topWatch2016$Freq),]
names(topWatch2016) = c("Brand","Freq")

#Merge to plot
timeseries = merge(topWatch2,topWatch2017, by.x = "Brand", by.y = "Brand")
timeseries = merge(timeseries,topWatch2016, by.x = "Brand", by.y = "Brand")
names(timeseries) = c("Brand","2018","2017","2016")
```

As an example, we can identify which watch brands have been picked up by the largest number of retailers since 2016. The below brands were picked up by more than 2 specialty retailers since 2016. 

```{r}
#Find which brands have been picked up by the most retailers
timeseries$Growth = (timeseries$"2018" - timeseries$"2016")
growtable = timeseries[order(-timeseries$Growth),]
knitr::kable(growtable[which(growtable$Growth > 2),c(1,4,2,5)],row.names = FALSE)
```

<br /> 

The below plot demonstrates the top 10 of the most carried brands by Specialty Retailers and their 3 year trend. 

```{r fig.height = 7, fig.width = 11}
#Only plot top30 from 2018
top30 = as.list(as.character(topWatch2$Brand[1:10]))
timeseries = timeseries[,1:4]
timeseries = timeseries[timeseries$Brand %in% top30,]

timeseriesPlot = melt(timeseries, id = "Brand")
timeseriesPlot$variable = as.numeric(as.character(timeseriesPlot$variable))


ggplot(timeseriesPlot, aes(x = variable, y = value, label = Brand)) + 
  geom_line(aes(color = Brand), size = 1) +
  ggtitle("3 Year Trend of Brands Being Carried\nBy Top 50 Specialty Retailers\n") +
  labs(caption = "Only top 10 carried brands of 2018 shown")+
  ylab("Number of Retailers Carrying Brand")+
  geom_label()+ 
  theme_wsj()+
  scale_x_continuous(breaks=seq(2016,2018,1))
```

<br /> 

### Conclusions: 

Drawing conclusions from and more generally maximizing the utility of this data / analysis, requires domain expertise. However, some quick observations can be gained by looking at this analysis: 

1. Up and coming brands: Bulova, G-shock, and Omega are all increasingly being carried by major retailers in the North American market. 

2. Top Brands: Citizen, Bulova, TAG Heuer, and Movado are consistently the most carried brands. 

3. Boutique Brands A Surprise: Surprisingly, smaller boutique brands seem to populate the top tier just below the top brands listed above. These include: 
    + Hamilton
    + Baume & Mercier
    + Longines

4. As expected Supersellers skew towards less-expensive brands: Supersellers to a much greater degree carry less elite brands in significant numbers such as: 
    + Timex
    + Casio
    + Seiko (unclear whether this includes Grand Seiko) 

5. The super-boutique watches are in the data, but not reflected in the top carried brands. Brands such as A. Lange & Söhne are likely to remain in the super-elite market. 
    